---
title: "Study 3, 5, 6 - causal perception - power simulation"
author: "Christoph Völter"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library("gghalves")

load("causal_perception_study3_5_6_power_sim_N40.RData")
```

## Generate data

```{r echo=FALSE, include=FALSE}
n.subject <- 40 # number subjects
n.per.subject <- 2 # observations per subject
n.per.condition <- 1 # observations per subject and condition
subj.id <- as.factor(paste("subj", str_pad(1:n.subject, 2, pad = "0"), sep = "."))
testgr.congruent.per <- 0 # test group performance in congruent condition
testgr.incongruent.per <- c(500) # test group performance in incongruent condition
controlgr.congruent.per <- 0 # control group performance in congruent condition
controlgr.incongruent.per <- 0 # control group performance in incongruent
```


```{r echo=FALSE, include=FALSE}
start.data <- data.frame(subj.id = rep(subj.id, times = 1, each = 2))
cb.data = data.frame(expand.grid(
  condition = c("congruent", "incongruent"),
  group = c("control", "test")
))
session = rep(c(1 , 2, 1, 2, 2, 1, 2, 1), times = n.subject / 4)

start.data <- cbind(start.data, cb.data, session)
start.data$sex <-
  as.factor(rep(x = c("f", "m", "m", "f"), times = n.subject / 4))[as.numeric(start.data$subj.id)]

# z-transformation of covariates
start.data$z.session <-
  as.vector(scale(as.numeric(start.data$session)))
```


# Data checks:

```{r echo=FALSE, include=FALSE}
# does each subject have only one sex and age?
xx <- table(start.data$subj.id, start.data$sex)
range(apply(X = xx > 0, MARGIN = 1, sum)) # should be 1 and 1

xx <- table(start.data$session, start.data$sex)
range(apply(X = xx > 0, MARGIN = 1, sum))

xx <- table(start.data$subj.id, start.data$condition)
range(apply(X = xx > 0, MARGIN = 1, sum))

xx <- table(start.data$condition, start.data$session)
range(apply(X = xx > 0, MARGIN = 1, sum))

xx <- table(start.data$subj.id, start.data$session)
range(apply(X = xx > 0, MARGIN = 1, sum))
```


## Simulation

### Test run

Preparation
```{r include=FALSE}
resid.sd <- 350#residual standard deviation
subj.sd <- 250#random effect of individual

icpt <- controlgr.congruent.per#intercept
sex.effect <- 0
session.effect <- 0

xdata <- start.data
tot.n <- nrow(xdata)#total n
m.mat <-
  model.matrix(object =  ~ group * condition + sex + z.session,
               data = xdata)

coefs <-
  c(
    "(Intercept)" = icpt,
    "grouptest" = testgr.congruent.per - icpt,
    "conditionincongruent" = controlgr.incongruent.per - icpt,
    "sexm" = sex.effect,
    "z.session" = session.effect,
    "grouptest:conditionincongruent" = (testgr.incongruent.per - testgr.congruent.per) - (controlgr.incongruent.per - controlgr.congruent.per)
  )

rv <- m.mat[, names(coefs)]%*%coefs+#fixed effects
rnorm(n=n.subject, sd=subj.sd)[as.numeric(subj.id)]+#random effect of subj.
rnorm(n=tot.n, sd=resid.sd)#residual variation
```


```{r }
#plotting
par(
  mar = c(2.7, 2.7, 0.2, 0.2),
  mgp = c(1.5, 0.3, 0),
  tcl = -0.15,
  las = 1,
  cex.lab = 0.7,
  cex.axis = 0.5
)
plot(interaction(xdata$condition, xdata$group), rv)
```


```{r }
library(lme4)
full = lmer(rv ~ group*condition + sex + z.session + (1 |
                                              subj.id),
            data = xdata,
            REML = F)
summary(full)$coef
drop1(full, test = "Chisq")
```

### Actual simulation
```{r eval=FALSE, include=FALSE}
n.simus <- 1000 # small number for testing
tot.n <- nrow(xdata)#total n
resid.sd <- 250#residual standard deviation
subj.sd <- 250#random effect of individual

icpt <- controlgr.congruent.per#intercept
sex.effect <- 0
order.effect <- 0

# create object to store the simulation parameters and results:
all.res <- data.frame(
  expand.grid(
    n.subject = n.subject,
    r.effect = subj.sd,
    resid.sd = resid.sd,
    testgr.congruent.per = testgr.congruent.per,
    testgr.incongruent.per = testgr.incongruent.per,
    controlgr.congruent.per = controlgr.congruent.per,
    controlgr.incongruent.per = controlgr.incongruent.per,
    simu = 1:n.simus
  )
)
all.res$icpt <- NA
all.res$grouptest <- NA
all.res$conditionincongruent <- NA
all.res$grouptest_conditionincongruent <- NA
all.res$re.sd <- NA
all.res$warns.full <- NA
all.res$warns.null <- NA
all.res$lrt.p.group_condition_int <- NA
all.res$lrt.p.sex <- NA
all.res$lrt.p.session <- NA
all.res$full.null.p <- NA

all.ests <- matrix(NA, nrow=n.simus, ncol=1)

# create data frame with design:
## done above

# load packages needed:
library(lme4)
# Loading required package: Matrix
library(kyotil) # we want to store info about convergence issues

# define control structure to make convergence more likely:
contr <- lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000))

xdata <- start.data

# run simulation
for (i in 1:nrow(all.res)) {
  set.seed(i) # allows to later replicate individual simulations
  
  m.mat <-
    model.matrix(object = ~ group * condition + sex + z.session,
                 data = xdata) # create model martix
  
  coefs <-
    c(
      "(Intercept)" = icpt,
      "grouptest" = all.res[i, "testgr.congruent.per"] - icpt,
      "conditionincongruent" = all.res[i, "controlgr.incongruent.per"] - icpt,
      "sexm" = sex.effect,
      "z.session" = session.effect,
      "grouptest:conditionincongruent" = (all.res[i, "testgr.incongruent.per"] -
                                            all.res[i, "testgr.congruent.per"]) - (all.res[i, "controlgr.incongruent.per"] -
                                                                                     all.res[i, "controlgr.congruent.per"])
    )
  # generate response:
  xdata$rv <- m.mat[, names(coefs)] %*% coefs + #fixed effects
    rnorm(n = n.subject, sd = subj.sd)[as.numeric(subj.id)] + #random effect of subj.
    rnorm(n = tot.n, sd = resid.sd)#residual variation
  
  # fit full model:
  full <-
    keepWarnings(lmer(rv ~ group*condition + sex + z.session + 
      (1 |subj.id),
      data = xdata,
      REML = FALSE,
      control = contr
    ))
  
  # fit null model:
  null <-
    keepWarnings(lmer(rv ~  sex + z.session + 
      (1 |subj.id),
      data = xdata,
      REML = FALSE,
      control = contr
    ))
  
  # store results:
  all.res[i, c("(Intercept)",
               "grouptest",
               "conditionincongruent",
               "sexm",
               "z.session",
               grouptest_conditionincongruent="grouptest:conditionincongruent")] <- fixef(full$value)
  all.res[i, "re.sd"] <-
    as.data.frame(summary(full$value)$varcor)[1, "sdcor"]
  all.res[i, "warns.full"] <-
    nchar(paste(full$warnings, collapse = ""))
  all.res[i, "warns.null"] <- nchar(paste(null$warnings, collapse = ""))
   all.res[i, "lrt.p.sex"] <-
    as.data.frame(drop1(full$value, test = "Chisq"))["sex", "Pr(Chi)"]
  all.res[i, "lrt.p.session"] <-
    as.data.frame(drop1(full$value, test = "Chisq"))["z.session", "Pr(Chi)"]
  all.res[i, "lrt.p.group_condition_int"] <-
    as.data.frame(drop1(full$value, test = "Chisq"))["group:condition", "Pr(Chi)"]
  all.res[i, "full.null.p"] <- as.data.frame(anova(null$value, full$value, test = "Chisq"))[2, "Pr(>Chisq)"]
}

save.image("causal_perception_study3_5_6_power_sim_N40.RData")
```

## Evaluation of results 

* number of warning per combinations of random effects (out of 1000 models per cell)  
Full model:  
```{r echo=FALSE}
#full model
tapply(X=all.res[, "warns.full"]>0, INDEX=all.res[, c("testgr.congruent.per", "testgr.incongruent.per", "controlgr.congruent.per", "controlgr.incongruent.per")],
FUN=sum)
#warning codes: 
#363: unable to evaluate scaled gradient. Model failed to converge: degenerate  Hessian with 1 negative eigenvalues
#205: Model is nearly unidentifiable: large eigenvalue ratio - Rescale variables?
```

* plotting the estimates (all models)

```{r echo=FALSE}
par(mar=c(3, 3, 0.2, 0.2), mgp=c(1.7, 0.3, 0), tcl=-0.15, las=1)
plot(
  x = as.numeric(as.factor(rep(
    x = c("(Intercept)", "grouptest_conditionincongruent", "re.sd"),
    each = nrow(all.res)
  ))),
  y = unlist(all.res[, c("(Intercept)", "grouptest:conditionincongruent", "re.sd")]),
  pch = 19, col = grey(level = 0.2, alpha = 0.2),
  xaxt = "n", xlim = c(0.5, 3.5), ylab = "estimate", xlab = ""
)
mtext(text = c("(Intercept)", "grouptest:conditionincongruent", "re.sd"), side = 1, at = 1:3, line = 0.2)
```

## Only models that converged are evaluated from here on:  

```{r include=FALSE}
all.res2=subset(all.res, warns.full==0)


```


### How many models converged, have a significant full-null model comparison, and a significant LRT of condition?  
```{r echo=FALSE}

lrt.data2 <- all.res2 %>%
  filter(full.null.p<0.05)%>%
  group_by(testgr.congruent.per, testgr.incongruent.per, controlgr.congruent.per, controlgr.incongruent.per) %>%
  summarise(lrt.p.int.mean2 = mean(lrt.p.group_condition_int), 
            n.sign.lrt_int = length(lrt.p.group_condition_int[lrt.p.group_condition_int < 0.05]), 
            n.lrt = n.simus,#length(lrt.p.con), 
            proportion.sign.lrt.int = length(lrt.p.group_condition_int[lrt.p.group_condition_int < 0.05]) / n.simus,
            proportion.sign.lrt.sex = length(lrt.p.group_condition_int[lrt.p.sex < 0.05]) / n.simus,
            proportion.sign.lrt.session = length(lrt.p.group_condition_int[lrt.p.session < 0.05]) / n.simus)

lrt.data2
```


