---
title: "Study 1 - causal perception - power simulation"
author: "Christoph Völter"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library("gghalves")

load("causal_perception_study1_power_sim_N24.RData")
```

## Generate data

```{r echo=FALSE, include=FALSE}
n.subject <- 24 # number subjects
n.per.subject <- 3 # observations per subject
n.per.condition <- 1 # observations per subject and condition
subj.id <- as.factor(paste("subj", str_pad(1:n.subject, 2, pad = "0"), sep = "."))

congruent.per <- 0 # performance in congruent condition
gap.per <- c(500) # performance in gap condition
lag.per <- c(500) # performance inlag condition
```

```{r echo=FALSE, include=FALSE}
# add condition and trial number
start.data <- data.frame(expand.grid(subj.id = subj.id, condition = c("congruent", "gap", "lag"), trial = c(1:n.per.condition))) %>%
  arrange(subj.id)

# add order
all_for_order <- expand.grid(p1 = 1:3, p2 = 1:3, p3 = 1:3, stringsAsFactors = FALSE) 
perms_order <- all_for_order[apply(all_for_order, 1, function(x) {length(unique(x)) == 3}),]
order<-c(unlist(perms_order[1,]), unlist(perms_order[2,]), unlist(perms_order[3,]),unlist(perms_order[4,]), unlist(perms_order[5,]), unlist(perms_order[6,]))
start.data$order <- as.factor(order)
start.data <- start.data %>% 
  arrange(subj.id, order)



# add demographics 
start.data$sex <- as.factor(rep(x = c("f", "m"), times = n.subject/2))[as.numeric(start.data$subj.id)]


# z-transformation of covariates
start.data$z.order <- as.vector(scale(as.numeric(start.data$order)))
```
# Data checks:
```{r echo=FALSE, include=FALSE}

# does each subject have only one sex and age?
xx <- table(start.data$subj.id, start.data$sex)
range(apply(X = xx > 0, MARGIN = 1, sum)) # should be 1 and 1

xx <- table(start.data$order, start.data$sex)
range(apply(X = xx > 0, MARGIN = 1, sum)) # should be 1 and 1

xx <- table(start.data$subj.id, start.data$condition)
range(apply(X = xx > 0, MARGIN = 1, sum))

xx <- table(start.data$condition, start.data$order)
range(apply(X = xx > 0, MARGIN = 1, sum))

xx <- table(start.data$subj.id, start.data$order)
range(apply(X = xx > 0, MARGIN = 1, sum))
```

## Simulation


Test
```{r eval=FALSE, include=FALSE}

resid.sd <- 350#residual standard deviation
subj.sd <- 500#random effect of individual
icpt <- congruent.per#intercept

sex.effect <- 0
order.effect <- 0

xdata<-start.data
tot.n <- nrow(xdata)#total n
m.mat <- model.matrix(object=~condition+sex+z.order, data <- xdata)
coefs <- c("(Intercept)"=icpt, "conditiongap"=gap.per, "conditionlag"=lag.per,  "sexm"=sex.effect, "z.order"=order.effect)

rv <- m.mat[, names(coefs)]%*%coefs+#fixed effects
rnorm(n=n.subject, sd <- subj.sd)[as.numeric(subj.id)]+#random effect of subj.
rnorm(n=tot.n, sd=resid.sd)#residual variation

#plotting
par(mar=c(2.7, 2.7, 0.2, 0.2), mgp=c(1.5, 0.3, 0),
tcl=-0.15, las=1, cex.lab=0.7, cex.axis=0.5)
plot(xdata$condition, rv)

library(lme4)
full=lmer(rv~condition+sex+z.order+(1|subj.id), data=xdata, REML=F)
summary(full)
drop1(full, test="Chisq")
```

Actual simulation

```{r eval=FALSE, include=FALSE}
n.simus <- 1000 # small number for testing
tot.n <- nrow(xdata)#total n
resid.sd <- 350#residual standard deviation
subj.sd <- 500#random effect of individual
icpt <- congruent.per#intercept
sex.effect <- 0
order.effect <- 0
age.effect <- 0

# create object to store the simulation parameters and results:
all.res <- data.frame(
  expand.grid(
    n.subject = n.subject,
    r.effect = subj.sd,
    resid.sd = resid.sd,
    gap.per = gap.per,
    lag.per = lag.per,
    congruent.per = congruent.per,
    simu = 1:n.simus
  )
)
all.res$icpt <- NA
all.res$conditiongap <- NA
all.res$conditionlag <- NA
all.res$re.sd <- NA
all.res$warns.full <- NA
all.res$lrt.p.con <- NA
all.res$lrt.p.sex <- NA
all.res$lrt.p.order <- NA

all.ests=matrix(NA, nrow=n.simus, ncol=1)
colnames(all.ests)=c("lrt.p.con")

# create data frame with design:
## done above

# load packages needed:
library(lme4)
# Loading required package: Matrix
library(kyotil) # we want to store info about convergence issues

# define control structure to make convergence more likely:
contr <-
  lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000))

xdata <- start.data

# run simulation
for (i in 1:nrow(all.res)) {
  set.seed(i) # allows to later replicate individual simulations
  
  m.mat <-
    model.matrix(object = ~ condition + sex + z.order, data = xdata) # create model martix
  
  coefs <- c(
    "(Intercept)" = congruent.per,
    "conditiongap" = gap.per,
    "conditionlag" = lag.per,
    "sexm" = 0,
    "z.order" = 0
  )
  # generate response:
  xdata$rv = m.mat[, names(coefs)] %*% coefs + #fixed effects
    rnorm(n = n.subject, sd = subj.sd)[as.numeric(subj.id)] + #random effect of subj.
    rnorm(n = tot.n, sd = resid.sd)#residual variation
  
  # fit full model:
  full <-
    keepWarnings(lmer(
      rv ~ condition + sex +  z.order + (1 | subj.id),
      data = xdata,
      REML = FALSE,
      control = contr
    ))

  # store results:
  all.res[i, c("(Intercept)",
               "conditiongap",
               "conditionlag",
               "sexm",
               "z.order")] <- fixef(full$value)
  all.res[i, "re.sd"] <-
    as.data.frame(summary(full$value)$varcor)[1, "sdcor"]
  all.res[i, "warns.full"] <-
    nchar(paste(full$warnings, collapse = ""))
    all.res[i, "lrt.p.con"] <-
    as.data.frame(drop1(full$value, test = "Chisq"))["condition", "Pr(Chi)"]
  all.res[i, "lrt.p.sex"] <-
    as.data.frame(drop1(full$value, test = "Chisq"))["sex", "Pr(Chi)"]
  all.res[i, "lrt.p.order"] <-
    as.data.frame(drop1(full$value, test = "Chisq"))["z.order", "Pr(Chi)"]
}

save.image("causal_perception_study1_power_sim_N24.RData")
```

## Evaluation of results

-   number of warning per combinations of random effects (out of 1000 models per cell)\
    Full model:\

```{r echo=FALSE}
#full model
tapply(X=all.res[, "warns.full"]>0, INDEX=all.res[, c("congruent.per", "gap.per", "lag.per")],
FUN=sum)
#warning codes: 
#363: unable to evaluate scaled gradient. Model failed to converge: degenerate  Hessian with 1 negative eigenvalues
#205: Model is nearly unidentifiable: large eigenvalue ratio - Rescale variables?
```



-   plotting the estimates (all models)

```{r echo=FALSE}
par(
  mar = c(3, 3, 0.2, 0.2),
  mgp = c(1.7, 0.3, 0),
  tcl = -0.15,
  las = 1
)
plot(
  x = as.numeric(as.factor(rep(
    x = c("(Intercept)", "conditiongap", "conditionlag", "re.sd"),
    each = nrow(all.res)
  ))),
  y = unlist(all.res[, c("(Intercept)", "conditiongap", "conditionlag", "re.sd")]),
  pch = 19,
  col = grey(level = 0.2, alpha = 0.2),
  xaxt = "n",
  xlim = c(0.5, 3.5),
  ylab = "estimate",
  xlab = ""
)
mtext(
  text = c("(Intercept)", "conditiongap", "conditionlag", "re.sd"),
  side = 1,
  at = 1:3,
  line = 0.2
)
```

## Only models that converged are evaluated from here on:

```{r include=FALSE}
all.res2=subset(all.res, warns.full==0)

```

### How many models converged, have a significant full-null model comparison, and a significant LRT of condition?

```{r echo=FALSE}

lrt.data2 <- all.res2 %>%
  group_by(gap.per, lag.per, congruent.per) %>%
  summarise(
    lrt.p.con.mean2 = mean(lrt.p.con),
    n.sign.lrt2 = length(lrt.p.con[lrt.p.con < 0.05]),
    n.lrt = n.simus,
    #length(lrt.p.con),
    proportion.sign.lrt2 = length(lrt.p.con[lrt.p.con < 0.05]) / n.simus
  )

lrt.data2
```
